\section{Results}
\label{sec:experiments}

This section reports the experimental evidence collected so far. The first result establishes a monolithic full-episode baseline. The later results characterise the subtask model and its performance inside the orchestrated execution loop.

\subsection{Full-Episode Imitation Baseline}

The first experiment evaluates whether changing the policy class is sufficient to improve the complete pick-and-place task when the task is treated as one monolithic episode. ACT and SmolVLA are trained on the same public HuggingFace dataset of 60 full-task demonstrations. In this setting, SmolVLA receives a single instruction for the whole episode, so the experiment does not evaluate meaningful language-conditioned control. Both policies are asked to solve the full task without any subtask-level monitoring.

\begin{table}[H]
    \centering
    \caption{Full-episode imitation results on the public dataset. Success means completing the full three-orange task.}
    \label{tab:full_episode_baseline}
    \begin{tabular}{lccc}
        \toprule
        Policy & Episodes & Success rate & Avg. oranges in plate \\
        \midrule
        ACT & 100 & 26\% & 1.62 / 3 \\
        SmolVLA & 100 & 28\% & 1.47 / 3 \\
        \bottomrule
    \end{tabular}
\end{table}

ACT and SmolVLA achieve comparable performance. This confirms that, under identical training conditions and with a single fixed instruction, the VLA model choice does not determine outcome. Selecting SmolVLA is therefore not a claim about its superiority as a policy class, but a choice motivated by its language conditioning capability --- which this experiment does not exercise and the subsequent subtask formulation is designed to exploit.

\subsection{Manual Subtask Dataset}

The dataset contains 921 successful subtask episodes, each annotated with a relative-position instruction and collected in a randomised pick order. The resulting SmolVLA model qualitatively follows target-specific grasp instructions — the enabling capability for the orchestrator: after a subtask failure, the system can issue a corrective instruction rather than aborting the full episode.

The model is not intended to solve the whole task as a monolithic policy. The formal metric of interest is its performance inside the orchestrated execution loop, where privileged simulator state is used to detect failed subtasks and trigger local retry or target redirection.

\begin{table}[H]
    \centering
    \caption{Orchestrated evaluation of the manually collected subtask model. The run was stopped after 326 of the planned 500 episodes. Success means completing the full three-orange task.}
    \label{tab:orchestrated_eval}
    \begin{tabular}{lccc}
        \toprule
        Model & Episodes & Success rate & Avg. oranges in plate \\
        \midrule
        Gal-pick-orange-tailedCH20 & 326 & 26/326 (7.98\%) & 1.32 / 3 \\
        \bottomrule
    \end{tabular}
\end{table}

The orchestrated result is worse than the monolithic baselines in Table~\ref{tab:full_episode_baseline}: its full-task success rate is 7.98\%, compared with 26\% for ACT and 28\% for the monolithic SmolVLA baseline, and its average number of oranges placed is also lower. The result therefore does not support the original expectation that subtask decomposition plus orchestration would immediately improve complete-task performance.

\begin{table}[H]
    \centering
    \caption{Subtask success rates during orchestrated evaluation, grouped by how many oranges were already in the plate at subtask start.}
    \label{tab:orchestrated_subtasks}
    \begin{tabular}{lccc}
        \toprule
        Subtask & 0 placed & 1 placed & 2 placed \\
        \midrule
        GRASP & 357/921 (38.8\%) & 204/608 (33.6\%) & 51/195 (26.2\%) \\
        LIFT & 302/357 (84.6\%) & 174/204 (85.3\%) & 39/51 (76.5\%) \\
        PLACE & 237/302 (78.5\%) & 140/174 (80.5\%) & 30/39 (76.9\%) \\
        \bottomrule
    \end{tabular}
\end{table}

The subtask breakdown shows that the main bottleneck remains GRASP. LIFT and PLACE are substantially more reliable once an orange has been acquired, but grasp success declines as the episode progresses and the scene moves further away from the original training distribution. The orchestrator also made heavy use of its recovery mechanisms: local retries occurred in 314 of 326 episodes, target redirections in 293 of 326 episodes, and 515 oranges were abandoned in total. These mechanisms prevented immediate dead-end execution, but they did not compensate for the low grasp reliability strongly enough to improve full-task success.

\subsection{Autonomous Data Generation}

Autonomous data generation is currently in progress. The orchestrated execution system is being used to collect successful subtask trajectories while filtering out failed subtask attempts. This is expected to produce a dataset that better reflects states encountered during autonomous execution than the manually teleoperated dataset alone.

The final results of this phase are pending. The following values should be reported once data generation and retraining are complete: FILLME: number of generated autonomous subtask episodes. FILLME: ratio of successful committed subtasks to failed attempts. FILLME: retraining result after autonomous data generation. FILLME: OOD evaluation protocol and result.
