\section{Methods}
\label{sec:methods}

This work studies orchestrated manipulation in a simulated pick-and-place task. The robot is an SO-101 arm operating in an Isaac Sim kitchen scene. Each episode contains three oranges and a plate. The objective is to place all three oranges into the plate. The scene provides privileged simulator state, which is used throughout to evaluate whether subtasks have achieved their intended physical outcome. This avoids confounding the orchestration problem with visual state-estimation errors.

\subsection{Full-Episode Imitation Baseline}

The first experimental setting treats the task as a single full-episode imitation-learning problem. A public HuggingFace dataset of 60 demonstrations is used, where each episode contains a complete sequence for picking the three oranges and placing them in the plate. The demonstrations are monolithic: they are not segmented into subtasks, and the learned policy is expected to execute the full task from the beginning of the episode to the end.

Two policy classes are compared: ACT, a lightweight diffusion-based model, and SmolVLA, a Vision-Language-Action model. Both are trained on the same dataset. SmolVLA receives a single fixed instruction for the whole episode, so neither model exploits language conditioning in any meaningful way. The comparison serves as a sanity check: if SmolVLA performs no better than ACT under identical conditions, it confirms that the VLA model choice alone does not determine performance in this setting.

\subsection{Subtask Decomposition}

Detecting and responding to intermediate failures requires knowing when one behaviour has ended and whether it succeeded. For this reason, the full task is reformulated as a sequence of shorter subtasks. The main subtasks are grasping a selected orange, lifting it, and placing it into the plate. These subtasks create explicit checkpoints where the execution system can evaluate whether the expected physical outcome has been achieved before committing to the next step.

In this formulation, a subtask failure is any case where the intended physical outcome is not reached within the allotted time. This includes failed grasps, an orange falling during lift, unsuccessful placement, or timeout.

An intermediate attempt was made by slicing the public full-episode dataset into shorter subtask segments. This produced a subtask-style dataset, but the original demonstrations always follow the same object order. As a result, the sliced data does not provide a meaningful signal connecting a language instruction to a specific target object: target identity is entangled with progress through the fixed sequence rather than with the instruction itself. This limitation motivated the creation of a custom dataset.

\subsection{Dataset Construction}

A custom teleoperated dataset was collected to support orchestrated execution. Unlike the public full-episode dataset, each episode in this dataset corresponds to one subtask rather than one complete three-orange trajectory. This makes it possible to discard failed teleoperation attempts, annotate each successful attempt with a subtask instruction, and train the policy on behaviours that can be sequenced by a higher-level controller.

The dataset contains instructions such as ``Grasp left orange'', ``Grasp middle orange'', ``Grasp right orange'', ``Pick it up'', and ``Place it into plate''. The pick order is randomised across episodes. This breaks the fixed-order correlation in the original demonstrations and makes relative-position instructions meaningful: the policy cannot infer the target from task progress alone and must instead learn to associate the instruction with the specified orange.

The final dataset contains 921 subtask episodes. The resulting SmolVLA model is trained as a subtask executor rather than as a monolithic full-task policy. Its role is to execute the current language-conditioned behaviour requested by the orchestrator.

\subsection{Orchestrator Design}

The VLA is responsible for three manipulation subtasks: GRASP(\textit{target}), LIFT, and PLACE. The orchestrator is responsible for subtask sequencing and outcome evaluation. It reads privileged simulator state to verify whether each subtask achieved its intended physical effect before advancing: it checks whether the gripper is holding the orange after GRASP, and whether the orange is in the plate after PLACE.

GRASP succeeds only when the selected orange is confirmed to be held according to geometric and force constraints. If the policy does not produce a valid grasp before the step limit, the subtask times out. The orchestrator then performs a scripted spatial reset to the home configuration and retries GRASP once for the same orange. This retry keeps the original sub-goal active and gives the policy a second attempt from a state where the target instruction is reliable.

LIFT and PLACE can fail either because the orange is dropped or because the subtask times out before reaching its intended outcome. PLACE also fails when the orange is not deposited in the plate. These failures trigger \textbf{local retry}: the orchestrator returns to GRASP for the same target orange without changing the language target. The original sub-goal is preserved, which is why local retry is the true recovery mechanism in this report.

After repeated failure on the same orange, the orchestrator uses \textbf{target redirection}. The current orange is abandoned and the next GRASP instruction targets a different available orange. This is not recovery of the original sub-goal: the system has given up on that orange for the remainder of the episode. However, it moves the execution loop out of a stuck state in which a monolithic policy could spend the remaining episode repeatedly attempting an unproductive grasp. Redirection therefore serves the global task objective by allowing progress on another orange.

Any change of language target is preceded by a \textbf{scripted spatial reset}. The VLA follows target-change instructions reliably only when the arm is far from the oranges; near a failed target, it tends to remain focused on the nearby object regardless of the new language instruction. The orchestrator therefore moves the arm to the home configuration with a deterministic joint-space interpolation before issuing a different GRASP target. This reset is a precondition for reliable language conditioning, not itself a recovery mechanism.

\subsection{Autonomous Data Generation}

The autonomous data generation phase used the orchestrated execution system to collect training data without manual teleoperation. This phase was motivated by the first formal evaluation of the handmade subtask model: subtasking plus orchestration did not improve full-task performance over the monolithic SmolVLA baseline. The next hypothesis was therefore that the subtask policy needed more data, especially data drawn from the states it visits during autonomous execution rather than only from manual demonstrations.

A full autonomous rollout can contain both failed and successful subtasks. Rather than discarding an entire rollout when a subtask fails, the system records the successful subtask trajectories and discards the failed ones.

This produces an automatically generated dataset of successful subtasks collected under the current policy's own execution distribution. Target redirection plays an important role here: without the ability to pivot to a different orange after a local failure, the system would need to discard the entire rollout. Redirection allows the system to continue collecting successful subtask data from scenes that would otherwise be wasted.

The autonomous generation run produced subtask episodes that were then filtered to balance the dataset by scene state: the number of episodes collected with one or two oranges already in the plate was equalised against those collected at the start of the episode, so the merged dataset does not over-represent early-scene behaviour. After filtering, 414 autonomous episodes were retained and merged with the 921 handmade episodes to form a final dataset of 1335 subtask episodes. A new model --- referred to as the merged subtask model --- was trained on this combined dataset.
