\section{Related Work}
\label{sec:related}

\textbf{Vision-language-action models.} Generalist manipulation policies increasingly couple a pretrained vision-language backbone to an action decoder trained on large cross-embodiment datasets: \mbox{RT-2}~\cite{brohan2023rt2} grounds web knowledge in control, Octo~\cite{octo2024} and OpenVLA~\cite{kim2024openvla} offer open generalist policies, and $\pi_0$~\cite{black2024pi0} introduces a flow-matching action expert. These models are large (billions of parameters) and assume substantial compute and data. SmolVLA~\cite{shukor2025smolvla}, used throughout this work, instead targets a single consumer GPU and community-scale data while inheriting the same flow-matching action-expert design. Strong single-skill competence, however, does not by itself guarantee reliable execution of a long-horizon task on a fixed low-cost arm --- the gap this work addresses.

\textbf{Imitation learning and its failure modes.} The policies above are trained by behavioural cloning, as are the action-chunking transformer ACT~\cite{zhao2023act} and diffusion policies~\cite{chi2023diffusion}. Cloning from successful demonstrations is prone to drift: small action errors carry the policy into states absent from the training distribution, where further mistakes compound --- the covariate-shift problem that motivated interactive correction such as DAgger~\cite{ross2011dagger}. Exactly how this error scales with task horizon is still actively debated~\cite{spencer2021feedback,foster2024horizon}, but the practical consequence is clear in our setting: a single fumbled grasp can leave the arm in a configuration the policy never recovers from. We take this as motivation to add structure \emph{around} the policy rather than to enlarge it.

\textbf{Decomposition and language hierarchies.} A common remedy is to break a long task into language-specified pieces. RT-H~\cite{belkhale2024rth} inserts a language-motion layer between task and action; SayCan~\cite{ahn2022saycan} grounds an instruction in a library of skills scored by affordance; and language-model planners such as Inner Monologue~\cite{huang2022inner} and Code-as-Policies~\cite{liang2022cap} sequence or generate skills from natural language. These systems delegate high-level control to a (often large) language model and typically assume each invoked skill succeeds. Our orchestrator keeps the decomposition into language-conditioned subtasks but replaces the language-model planner with a lightweight finite-state controller that verifies each subtask's physical outcome from privileged state, making intermediate failure observable rather than assumed away.

\textbf{Failure detection and recovery.} Recent work detects and reacts to execution failures: RACER~\cite{dai2024racer} augments demonstrations with recovery trajectories supervised by a vision-language model; SAFE~\cite{gu2025safe} learns a task-agnostic failure detector for VLAs; Code-as-Monitor~\cite{zhou2024codeasmonitor} and Sentinel~\cite{agia2024sentinel} monitor execution at runtime; and REFLECT~\cite{liu2023reflect} uses a language model to explain failures for replanning. These approaches either require failure-labelled data or add a learned/language-model monitor on top of the policy. We instead define recovery \emph{structurally}, at two explicit levels of the goal hierarchy --- retrying the same target (Level-1) versus redirecting to a different feasible target to preserve task progress (Level-2) --- adjudicated directly by privileged physical checks. Crucially, none of these works address a precondition we find essential: language-based target switching is reliable only from a spatially reset configuration (\cref{sec:methods}), a constraint specific to language-conditioned VLAs that motivates our scripted reset.

\textbf{Self-improvement from autonomous data.} Finally, our autonomous data-generation loop relates to policy self-improvement: keeping only the successful self-collected trajectories and retraining is a reward-filtered form of self-imitation --- the binary-reward limit of reward-weighted regression~\cite{peters2007rwr} --- and is related to autonomous-practice systems such as MEDAL++~\cite{sharma2023selfimproving}. Here the orchestrator's privileged success checks supply the filter; \cref{sec:experiments} reports when this loop helps and when it does not.
