\section{Discussion}
\label{sec:discussion}

This section interprets the results against the central hypothesis --- that wrapping the policy in execution-level recovery makes the long-horizon task more robust --- tracing what decomposition achieved, why grasp acquisition remains the binding constraint, and why neither autonomous data nor cheaper fine-tuning tweaks closed the gap, before turning to the conclusion, limitations, and future work.

\textbf{Decomposition broadened partial progress, not completion.} The hypothesis (\cref{sec:introduction}) was that wrapping the policy in subtask checkpoints with two-level recovery would make the long-horizon task more robust. The realised payoff is narrower: the orchestrated system almost never fails \emph{completely}, placing at least one orange in 97\% of Teleop episodes and 84\% on Teleop+Auto, against 56\% and 64\% for the standard monotask policy. The mechanism is \textbf{Level-2} redirection --- giving up on an ungraspable orange to try another: the orchestrator redirects after every grasp stall, while the monotask policy switches away in only about half of them --- and nothing ties its switch to the failure: trained on perfect demonstrations, where no grasp fails and no target is ever abandoned, it has never seen redirection as a response to failure, so whether it stays or leaves is closer to a coin flip than a decision. The orchestrated recoveries come almost entirely through a different orange (53\% vs.\ 16\% of Teleop stalls, \cref{sec:failure_modes}). It did not reliably raise full-task completion under the shared standard recipe: the subtask formulation gains six points on Teleop (14\% to 20\%) but loses five on Teleop+Auto (18\% to 13\%, mean occupancy still ahead) --- small and inconsistent next to the near-elimination of empty episodes. Under full fine-tuning, though, the subtask formulation edges ahead on completion on both sources while the empty-episode gap nearly closes (\cref{sec:experiments}): the structure matters most while grasping is weak. Under the standard recipe, decomposition secures more oranges, not more finished tasks.

\textbf{The grasp is the only hard step, and decomposition does not make it easier.} Once an orange is held the rest is reliable: for the standard models, lift and place each succeed 80--92\% of the time, and a secured orange reaches the plate 73--87\% of the time (the fully-tuned models sit higher still). The limiting step is acquisition --- only about 37\% of grasp attempts secure an orange (55\% for the fully-tuned Teleop subtask model), and the policy grasps the \emph{requested} orange just 64--73\% of the time, worst on \emph{right} (\cref{sec:obedience}). The orchestrator's checkpoints sit at boundaries that are already dependable; no amount of monitoring places an orange the policy never grasps. The three-way GRASP/LIFT/PLACE split is thus finer than the failure structure warrants: \textbf{Level-1} local retry almost never decides an outcome --- only about 6--11\% of placed oranges took more than one grasp --- so the realised contribution collapses onto Level-2 redirection alone, itself capped by how often a reachable alternative remains.

\textbf{Autonomous data never helped the subtask model --- and helped the monotask only under the weaker recipes.} Adding 414 self-generated successful subtasks \emph{helped} the standard monotask policy (full-task 14\% to 18\%, mean 1.02 to 1.29, empty episodes 44\% to 36\%) yet \emph{regressed} the orchestrated one (20\% to 13\%, mean 1.77 to 1.42, secured oranges per episode 2.13 to 1.94, a held orange placed 87\% down to 73\%). Why is left open. One reading is that subtasking's advantage is redirection, which pays off only when the policy gets stuck: if the extra data made grasping more reliable, redirection would matter less and the formulation's minor overheads would dominate. Two observations keep this unresolved. First, the standard monotask's secured-per-episode rises on the merged data (1.26 to 1.47), weakly consistent with ``gets stuck less'' --- but the fully-tuned pair reverses this, and at that strength the merged data hurts \emph{both} formulations (\cref{sec:experiments}). Second, the orchestrated policy still fails completely far less often on Teleop+Auto (16\% vs.\ 36\% empty), so redirection is plainly still active --- arguing against the idea that it simply lost its purpose.

\textbf{More self-generated data is unlikely to be the fix.} The autonomous loop is a weak improvement mechanism by construction. SmolVLA produces each action chunk by integrating a flow-matching field from a Gaussian noise sample (\cref{subsec:finetuning}); trained by behavioural cloning on near-deterministic teleoperation, it largely reproduces the demonstrated motion rather than exploring alternative grasps. The harvested episodes are correspondingly narrow --- a self-distribution that reinforces what the policy already does well. The loop also assigns a single success bit to a whole subtask with no within-chunk credit, a reward-weighted behavioural-cloning round rather than genuine policy search (\cref{subsec:autonomous}). A near-deterministic policy filtered by a coarse binary reward cannot discover the grasp it is missing; closing the acquisition gap likely needs explicit exploration --- for instance online reinforcement fine-tuning of the flow policy --- or targeted demonstrations on the hard \emph{right} and late-scene grasps, not more in-distribution data.

\textbf{Cheap fine-tuning tweaks give no reliable gain; the full recipe does.} The partial recipe (\cref{subsec:finetuning}) clearly helps only the Teleop subtask model (full-task 20\% to 33\%, mean 1.77 to 1.88); on the other three formulations the change is mixed and within sampling noise (Teleop monotask 14\% to 9\%, Teleop+Auto subtask 13\% to 10\%, Teleop+Auto monotask 18\% to 25\%). Removing the terminal freeze frames (\cref{subsec:teleop}), tried on the Teleop subtask alone, reaches 32\%/1.81 --- essentially the partial-tuning result from a pure data change, so that gain is not clearly an effect of the extra trainable weights. Because each action chunk is sampled from a fresh noise seed, differences of a few points over 100 episodes may reflect sampling noise rather than a real ordering. The one decisive change is full fine-tuning: unfreezing everything, vision encoder included, improves \emph{every} family by 18 to 40 points (\cref{sec:experiments}), on the Teleop subtask model by raising per-attempt grasp success from 37\% to 55\%. Its optimisation regime differs, but the regime control bounds that contribution as minor and the LightwheelAI comparison shares one regime (\cref{subsec:limitations}). Notably, the gain is visuomotor, not linguistic: the fully-tuned models obey their target instruction no better than the others (\cref{sec:obedience}).

\subsection{Conclusion}

This project asked whether a pick-and-place robot becomes more reliable when a controller is wrapped around its learned policy. The task is split into simple steps --- grasp an object, lift it, place it --- and after each step the controller checks whether it actually worked; on failure the robot retries the same object or gives up and turns to another, so the task can still make progress. The payoff is real but selective. At the subgoal level the orchestration clearly helps --- most while the policy's grasping is still weak: the robot almost never ends empty-handed, because it can abandon an object it cannot pick up and still place a different one. Full completion of this task, it did not reliably raise: the hardest step --- grasping the object the robot is asked for --- comes before any check or recovery can help, and what moved that bottleneck was a stronger policy (fine-tuning the whole network, vision included), not more structure. The two levers are complementary: better grasping raises how far each attempt gets, while the orchestration makes failure gradual instead of total. Moreover, this task leaves a controller little to decide --- three interchangeable objects, one ordering-free goal --- so the orchestration shows its floor here rather than its ceiling; on longer tasks with distinct objects, ordering constraints, and irreversible steps, the same checkpoint-and-recover structure has more room to contribute.

\subsection{Limitations}
\label{subsec:limitations}

Several limitations qualify these results. The broadest is interpretive: the system has many interacting design choices --- dataset composition, task formulation, which layers are fine-tuned, the per-subtask step budgets --- and only some are varied here, so the comparisons are best read as trends rather than as isolating a single decisive factor. Teasing each effect apart would take more experiments than this study runs.

\textbf{Privileged supervision.} Subtask outcomes are read from privileged simulator state, which a real deployment would have to replace with perception; the monotask grasp trace is reconstructed offline from that state, so its grasp attempts cannot be counted on the same footing as the orchestrated ones and no per-attempt monotask grasp \emph{rate} is reported.

\textbf{Limited sampling.} Each evaluation covers only 100 seeded episodes, so small differences between models may reflect sampling noise rather than a real effect.

\textbf{An unoptimised orchestrator.} The per-subtask step budgets and the finite-state machine itself are hand-set heuristics rather than optimised choices.

\textbf{Inconsistent composite-label boundaries.} The composite spatial labels (\emph{top right}, \emph{bottom right}) apply only when two oranges are laterally too close to separate, and the boundary was drawn differently across sources: teleop task strings were typed by hand, judging closeness by eye, while the autonomous data and the evaluation apply a scripted 3\,cm threshold. The conventions can therefore disagree both \emph{within} the merged training set and between training and evaluation --- a near-aligned pair the rule calls \emph{top right}/\emph{bottom right} may have been annotated \emph{right}/\emph{middle} by eye, or vice versa --- leaving composite-label supervision scarce (16 grasp demonstrations for \emph{bottom right} across both sources, against 209 for \emph{right}) and of uncertain consistency. The evaluation-side evidence fits a supervision gap rather than genuinely ambiguous scenes: the bottom-right mixups (\cref{sec:obedience}) do not cluster near the 3\,cm boundary (obeyed and misgrabbed requests share the same lateral-gap distribution, median 2.1\,cm) and the close pair is always clearly separated on the depth axis ($\geq$5\,cm). How often the hand-typed labels actually disagree with the rule was not audited, and none of this explains the dominant leftward bias.

\textbf{An imperfectly matched fully-tuned regime.} Training all \num{450}M parameters did not fit in GPU memory at the shared batch size, so the fully-tuned models trained with batch size 32 instead of 64 and \num{40000} gradient steps instead of \num{20000} --- the same number of samples seen, but twice as many weight updates --- and kept the pretrained checkpoint's action-chunk length of 50 rather than the custom models' 20. Two checks bound this confound without eliminating it: the regime control (the partially-tuned Teleop subtask model retrained under the full regime, chunk length included) gains 7 points where full tuning gains 25, and the LightwheelAI Baseline's standard model already used the same batch, steps, and chunk length, so that family's 20-point gain involves no regime change at all. On the remaining families the regime's exact share is not isolated.

\textbf{Slow execution.} The teleoperated demonstrations were deliberately slow --- precise gamepad control demanded it (\cref{subsec:teleop}) --- so the policy inherited slow motions and each subtask consumes many simulation steps, which makes evaluation rollouts slow and the approach less suited to time-critical use.

\textbf{Chronologically constrained subtask data.} The subtasks were collected as isolated fragments in their natural chronological order, never starting a subtask from an out-of-sequence state. This was deliberate, so that the \emph{same} trajectories could serve the matched monotask comparison (\cref{subsec:monotask}), which is only meaningful for chronologically ordered steps --- being on one orange and then leaving it for another is senseless end-to-end. The effect is a policy whose behaviour stays entangled with apparent task progress rather than driven purely by the instruction: it follows the prompt only about two-thirds of the time (\cref{sec:obedience}) and cannot redirect from a state near an orange on its own, which is precisely why the scripted home-pose reset is required (\cref{subsec:failure_taxonomy}). Training on isolated fragments rather than complete, imperfect trajectories of the whole task also plausibly limits long-horizon chaining and off-distribution robustness. How a from-any-state dataset would lift these constraints is taken up in \cref{subsec:future_subtask}.

\subsection{Future Work}

The two levels of the system suggest two largely independent directions: making the atomistic subtask policy more capable, and generalising the orchestration around it.

\subsubsection{Subtask Execution}
\label{subsec:future_subtask}

The binding constraint --- grasp acquisition --- lies within the atomistic subtask policy, so making that policy more capable is the priority, and this is largely a matter of data. More demonstrations and higher-quality ones would help directly: the SO-101's five actuated arm joints are one short of the six degrees of freedom a full end-effector pose requires, so the arm cannot track a full hand pose --- a Meta Quest 3 hand-tracking pipeline built late in the project proved less intuitive than the gamepad for this reason and collected no data. A 6-DOF arm would make teleoperation natural enough to collect a larger, more diverse, and cleaner dataset.

Quality also means \emph{coverage}. The chronologically-ordered data never exercises a subtask from an out-of-sequence state (\cref{subsec:limitations}); deliberately collecting such episodes --- being asked to grasp the right orange while already holding or hovering over another --- would make the policy a true from-any-state subtask executor that obeys the instruction from any configuration, and would let it redirect on its own without the scripted home-pose reset.

\subsubsection{Orchestration}

The high-level orchestration is largely orthogonal to the policy and can generalise along two axes. First, it currently reads each subtask's outcome from privileged simulator state, which is unavailable on a real robot (\cref{subsec:failure_taxonomy}); judging success from observations alone --- for instance with a vision-language model (VLM) as a success monitor --- would make the same control structure deployable in reality.

Second, the value of an explicit orchestrator should grow with task complexity. The present task is simple enough that a hand-designed finite-state machine suffices, but for long-horizon tasks with many object types, ordering constraints, and irreversible steps, enumerating and demonstrating every case end-to-end becomes infeasible --- exactly where a high-level controller earns its place. Given a reliable library of atomistic subtask policies, the decision of which subtask to run and when could be delegated either to a finite-state machine, when the task structure is known, or to an external language or vision-language model, when it is not, with the VLA optionally signalling subtask completion.
